`You Draw It': Implementation of visually fitted trends with \texttt{r2d3}

### 2022-07-27 2022-sdss-you-draw-it-manuscript.tex
added author information

\usepackage[dvipsnames]{xcolor} % colors
\newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
\newcommand{\rh}[1]{{\textcolor{Green}{#1}}}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={`You Draw It': Implementation of visually fitted trends with r2d3},
  pdfkeywords={graphics, user interaction, regression},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{`You Draw It': Implementation of visually fitted trends with
\texttt{r2d3}}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
How do statistical regression results compare to intuitive, visually
fitted results? Fitting lines by eye through a set of points has been
explored since the 20th century. Common methods of fitting trends by eye
involve maneuvering a string, black thread, or ruler until the fit is
suitable, then drawing the line through the set of points. In 2015, the
New York Times introduced an interactive feature, called `You Draw It,'
where readers are asked to input their own assumptions about various
metrics and compare how these assumptions relate to reality. This
research is intended to implement `You Draw It', adapted from the New
York Times, as a way to measure the patterns we see in data.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi

@@

We set out to implement `You Draw It', adapted from the NYT feature, as
a way to experimentally assess the patterns we see in data. Here, we
provide technical details of the software development, utilizing
interactive graphics in R (R Core Team 2022). We then share results from
our study which validates `You Draw It' as a method for graphical
testing on linear trends and apply an appropriate data analysis method
to the participant data. We also briefly demonstrate the use of the `You
Draw It' method and analysis on nonlinear data.

@@

\hypertarget{extension-to-nonlinear-data}{%
\subsection{Extension to Nonlinear
Data}\label{extension-to-nonlinear-data}}

The analysis method presented above extends nicely beyond the linear
regressions tested in Mosteller et al. (1981). Here we briefly
demonstrate this with an example from a study examining participants
ability to make forecasts for exponentially increasing data on log and
linear scales. Along with analyzing the feedback data with the GAMM
method for flexibility due to nonlinear data, we used spaghetti plots to
conduct a visual analysis of participant forecasts compared to the
nonlinear least squares statistical model and make comparisons between
two chart design features Figure~\ref{fig-exponential-spaghetti-plot}.
The combination of the GAMM analysis and visual display demonstrates the
strength of the `You Draw It' method for testing statistical graphics.

@@

\begin{itemize}
\tightlist
\item
  \textbf{`You Draw It' Demonstration Applet:} The shiny app used to
  demonstrate the `You Draw It' method can be accessed at
  \href{https://emily-robinson.shinyapps.io/you-draw-it-validation-applet/}{emily-robinson.shinyapps.io/you-draw-it-validation-applet}.
\item
  \textbf{`You Draw It' Script:} The `You Draw It' JavaScript code can
  be accessed at
  \href{https://github.com/earobinson95/presentations/blob/master/can-you-draw-it/www/you-draw-it.js}{github.com/earobinson95/presentations/blob/master/can-you-draw-it/www/you-draw-it.js}.
\item
  \textbf{Study Applet:} The shiny app used to conduct the study can be
  accessed at
  \href{https://shiny.srvanderplas.com/perception-of-statistical-graphics/}{shiny.srvanderplas.com/perception-of-statistical-graphics/}.
\item
  \textbf{Study Applet Code:} The code used to create the RShiny Applet
  for data collection can be found at
  \href{https://github.com/earobinson95/log-perception-prolific/tree/main/perception-of-statistical-graphics}{github.com/earobinson95/log-perception-prolific/tree/main/perception-of-statistical-graphics}.
\item
  \textbf{Participant Data (Linear):} De-identified participant data
  collected in the study and used for analyses are available to be
  downloaded from GitHub at
  \href{https://github.com/earobinson95/sdss-2022-you-draw-it-manuscript/raw/master/data/eyefitting-model-data.csv}{github.com/earobinson95/sdss-2022-you-draw-it-manuscript/raw/master/data/eyefitting-model-data.csv}.
\item
  \textbf{Participant Data (Nonlinear):} De-identified participant data
  collected in the study and used for analyses are available to be
  downloaded from GitHub at
  \href{https://github.com/earobinson95/sdss-2022-you-draw-it-manuscript/raw/master/data/youdrawit-model-data.csv}{github.com/earobinson95/sdss-2022-you-draw-it-manuscript/raw/master/data/youdrawit-model-data.csv}.
\item
  \textbf{Data Analysis Code:} The code used to replicate the nonlinear
  study analysis in this paper can be found at
  \href{https://earobinson95.github.io/sdss-2022-you-draw-it-manuscript/analysis/data-analysis.html}{earobinson95.github.io/sdss-2022-you-draw-it-manuscript/analysis/data-analysis.htmll}.
\end{itemize}

>>>>

\author[1]{Emily A. Robinson\thanks{Corresponding author. Email: erobin17@calpoly.edu}}
\author[2]{Reka Howard\footnote{Email: rekahoward@unl.edu}}
\author[2]{Susan VanderPlas\footnote{Email: susan.vanderplas@unl.edu}}
\affil[1]{Department of Statistics, CalPoly-San Luis Obispo}
\affil[2]{Department of Statistics, University of Nebraska-Lincoln}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={`You Draw It': Implementation of visually fitted trends with r2d3},
  pdfkeywords={graphics, user interaction, regression},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{`You Draw It': Implementation of visually fitted trends with
\texttt{r2d3}}
% \author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
How do statistical regression results compare to intuitive, visually
fitted results? Fitting lines by eye through a set of points has been
explored since the 20th century. Common methods of fitting trends by eye
involve maneuvering a string, black thread, or ruler until the fit is
suitable, then drawing the line through the set of points. In 2015, the
New York Times introduced an interactive feature, called `You Draw It,'
where readers are asked to input their own assumptions about various
metrics and compare how these assumptions relate to reality. This
research is intended to implement `You Draw It', adapted from the New
York Times, as a way to measure the patterns we see in data. In this
paper, we describe the adaptation of an old tool for graphical testing
and evaluation, eye-fitting, for use in modern web-applications suitable
for testing statistical graphics. We present an empirical evaluation of
this testing method for linear regression, and briefly discuss an
extension of this method to non-linear applications.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, breakable, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

We set out to implement `You Draw It', adapted from the NYT feature, as
a way to experimentally assess the patterns we see in data. Here, we
provide technical details of the software development, utilizing
interactive graphics in R (R Core Team 2022). We then share results from
our study which validates `You Draw It' as a method for graphical
testing on linear trends and apply an appropriate data analysis method
to the participant data. We also briefly demonstrate the use of the `You
Draw It' method and analysis on non-linear data.

@@

\hypertarget{extension-to-non-linear-data}{%
\subsection{Extension to Non-linear
Data}\label{extension-to-non-linear-data}}

The analysis method presented above extends nicely beyond the linear
regressions tested in Mosteller et al. (1981). Here we briefly
demonstrate this with an example from a study examining participants
ability to make forecasts for exponentially increasing data on log and
linear scales. Along with analyzing the feedback data with the GAMM
method for flexibility due to non-linear data, we used spaghetti plots
to conduct a visual analysis of participant forecasts compared to the
non-linear least squares statistical model and make comparisons between
two chart design features Figure~\ref{fig-exponential-spaghetti-plot}.
The combination of the GAMM analysis and visual display demonstrates the
strength of the `You Draw It' method for testing statistical graphics.

@@

\begin{itemize}
\tightlist
\item
  \textbf{`You Draw It' Demonstration Applet:} The shiny app used to
  demonstrate the `You Draw It' method can be accessed at
  \href{https://emily-robinson.shinyapps.io/you-draw-it-validation-applet/}{emily-robinson.shinyapps.io/you-draw-it-validation-applet}.
\item
  \textbf{`You Draw It' Script:} The `You Draw It' JavaScript code can
  be accessed at
  \href{https://github.com/earobinson95/presentations/blob/master/can-you-draw-it/www/you-draw-it.js}{github.com/earobinson95/presentations/blob/master/can-you-draw-it/www/you-draw-it.js}.
\item
  \textbf{Study Applet:} The RShiny applet used to conduct the linear
  and non-linear studies can be accessed at
  \href{https://shiny.srvanderplas.com/perception-of-statistical-graphics/}{shiny.srvanderplas.com/perception-of-statistical-graphics/}.
\item
  \textbf{Study Applet Code:} The code used to create the RShiny applet
  for data collection can be found at
  \href{https://github.com/earobinson95/log-perception-prolific/tree/main/perception-of-statistical-graphics}{github.com/earobinson95/log-perception-prolific/tree/main/perception-of-statistical-graphics}.
\item
  \textbf{Participant Data (Linear):} De-identified participant data
  collected in the linear study and used for analyses are available to
  be downloaded from GitHub at
  \href{https://github.com/earobinson95/sdss-2022-you-draw-it-manuscript/raw/master/data/eyefitting-model-data.csv}{github.com/earobinson95/sdss-2022-you-draw-it-manuscript/raw/master/data/eyefitting-model-data.csv}.
\item
  \textbf{Participant Data (Non-linear):} De-identified participant data
  collected in the non-linear study and used for analyses are available
  to be downloaded from GitHub at
  \href{https://github.com/earobinson95/sdss-2022-you-draw-it-manuscript/raw/master/data/youdrawit-model-data.csv}{github.com/earobinson95/sdss-2022-you-draw-it-manuscript/raw/master/data/youdrawit-model-data.csv}.
\item
  \textbf{Data Analysis Code:} The code used to replicate the linear and
  non-linear study analysis in this paper can be found at
  \href{https://earobinson95.github.io/sdss-2022-you-draw-it-manuscript/analysis/data-analysis.html}{earobinson95.github.io/sdss-2022-you-draw-it-manuscript/analysis/data-analysis.htmll}.
\end{itemize}
